---
title: "PM2.5 levels and Subjective cognitive decline"
author: "Aakash Parekh and Dr. Erik Parker"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 1
    code_folding: hide
    df_print: kable
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(zoo)
library(mice)
library(lmerTest)
library(gtsummary)
library(sjPlot)
library(patchwork)

# ── Read in cognitive decline data ────────────────────────────────────────────
Total_Cog_Decline <- read.csv("~/Desktop/CH internship/New versions BRFSS data/cognitive_decline_by_state_2015_to_2023.csv")
# Pivot from wide to long
Total_Cog_Decline_long <- Total_Cog_Decline %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "Year",
    names_prefix = "X",
    values_to = "MeanCognitiveDecline"
  ) %>%
  mutate(Year = as.integer(Year))

# Count non-NA data points per state
state_data_counts <- Total_Cog_Decline_long %>%
  group_by(StateName) %>%
  summarise(years_with_data = sum(!is.na(MeanCognitiveDecline)))

# Filter out states with 2 or fewer years of data
states_with_enough_data <- state_data_counts %>%
  filter(years_with_data > 2) %>%
  pull(StateName)

Total_Cog_Decline_long_filtered <- Total_Cog_Decline_long %>%
  filter(StateName %in% states_with_enough_data)

# Linear interpolation between yearly points
Total_Cog_Decline_long_interp <- Total_Cog_Decline_long %>%
  group_by(StateName) %>%
  arrange(Year) %>%
  mutate(InterpolatedLine = na.approx(MeanCognitiveDecline, Year, na.rm = FALSE)) %>%
  ungroup()

# ── Read in PM2.5 data ────────────────────────────────────────────────────────
pm25_all_states_years <- read_csv("~/Desktop/CH internship/New versions BRFSS data/pm25_all_states_years.csv") %>%
  select(-State.Code)

# Calculate lags for PM2.5
pm25_all_states_years <- pm25_all_states_years %>%
  arrange(StateName, Year) %>%
  group_by(StateName) %>%
  mutate(
    PM25_lag1 = lag(Mean_PM25, n = 1),
    PM25_lag2 = lag(Mean_PM25, n = 2),
    PM25_lag3 = lag(Mean_PM25, n = 3)
  ) %>%
  ungroup()

# ── Read in confounder data ───────────────────────────────────────────────────
state_confounders <- read_csv("~/Desktop/CH internship/New versions BRFSS data/state_confounders_2015_2023.csv")

# ── Merge all data ────────────────────────────────────────────────────────────

# Merge PM2.5 and confounders with full interpolated cognitive decline data
merged_long_data <- Total_Cog_Decline_long_interp %>%
  left_join(pm25_all_states_years, by = c("StateName", "Year")) %>%
  left_join(state_confounders, by = c("StateName", "Year"))

# Merge PM2.5 and confounders with filtered cognitive decline data (3+ obs states only)
filtered_merged_long_data <- Total_Cog_Decline_long_filtered %>%
  left_join(pm25_all_states_years, by = c("StateName", "Year")) %>%
  left_join(state_confounders, by = c("StateName", "Year"))

# ── Multiple imputation ───────────────────────────────────────────────────────

# Pivot to wide format for imputation
wide_merged_data <- merged_long_data %>%
  select(-InterpolatedLine) %>%
  pivot_wider(
    values_from = c("MeanCognitiveDecline", "Mean_PM25", "PM25_lag1", "PM25_lag2", "PM25_lag3",
                    "pct_age65plus", "pct_depression", "pct_smoker", "pct_stroke",
                    "pct_lowedu", "pct_lowincome", "pct_nonwhite"),
    names_from = "Year"
  )

# Run mice imputation
imputed_data <- mice(wide_merged_data,
                     m = 10,
                     maxit = 10,
                     method = "pmm",
                     seed = 8675309,
                     verbose = FALSE)

# Extract long format from imputed data (exclude = FALSE keeps original unimputed dataset)
imputed_complete <- complete(imputed_data, action = "long", include = FALSE)

# Pivot each variable back to long format
cog_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("MeanCognitiveDecline_"),
               names_to = "Year", names_prefix = "MeanCognitiveDecline_",
               values_to = "MeanCognitiveDecline")

pm25_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("Mean_PM25_"),
               names_to = "Year", names_prefix = "Mean_PM25_",
               values_to = "Mean_PM25")

pm25_lag1_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("PM25_lag1_"),
               names_to = "Year", names_prefix = "PM25_lag1_",
               values_to = "PM25_lag1")

pm25_lag2_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("PM25_lag2_"),
               names_to = "Year", names_prefix = "PM25_lag2_",
               values_to = "PM25_lag2")

pm25_lag3_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("PM25_lag3_"),
               names_to = "Year", names_prefix = "PM25_lag3_",
               values_to = "PM25_lag3")

conf_age_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_age65plus_"),
               names_to = "Year", names_prefix = "pct_age65plus_",
               values_to = "pct_age65plus")

conf_dep_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_depression_"),
               names_to = "Year", names_prefix = "pct_depression_",
               values_to = "pct_depression")

conf_smok_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_smoker_"),
               names_to = "Year", names_prefix = "pct_smoker_",
               values_to = "pct_smoker")

conf_stroke_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_stroke_"),
               names_to = "Year", names_prefix = "pct_stroke_",
               values_to = "pct_stroke")

conf_edu_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_lowedu_"),
               names_to = "Year", names_prefix = "pct_lowedu_",
               values_to = "pct_lowedu")

conf_inc_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_lowincome_"),
               names_to = "Year", names_prefix = "pct_lowincome_",
               values_to = "pct_lowincome")

conf_race_long <- imputed_complete %>%
  pivot_longer(cols = starts_with("pct_nonwhite_"),
               names_to = "Year", names_prefix = "pct_nonwhite_",
               values_to = "pct_nonwhite")

# Join all long datasets together
imputed_long <- cog_long %>%
  left_join(pm25_long,       by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag1_long,  by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag2_long,  by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag3_long,  by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_age_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_dep_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_smok_long,  by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_stroke_long,by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_edu_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_inc_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_race_long,  by = c("StateName", "Year", ".imp", ".id")) %>%
  select(.imp, .id, StateName, Year, MeanCognitiveDecline, Mean_PM25,
         PM25_lag1, PM25_lag2, PM25_lag3,
         pct_age65plus, pct_depression, pct_smoker, pct_stroke,
         pct_lowedu, pct_lowincome, pct_nonwhite)

```


# Descriptive Statistics {.tabset}

## Average cognitive decline by state (raw data)

```{r}

Total_Cog_Decline_long %>%
  group_by(StateName) %>%
  summarise(Average_Decline = mean(MeanCognitiveDecline, na.rm = TRUE)) %>%
  arrange(desc(Average_Decline)) %>%
  print(n = 52)

```

## Average cognitive decline by state (imputed data)

```{r}

imputed_long %>%
  group_by(StateName) %>%
  summarise(Average_Decline = mean(MeanCognitiveDecline, na.rm = TRUE)) %>%
  arrange(desc(Average_Decline)) %>%
  print(n = 52)

```

## Average PM2.5 by state

```{r}

pm25_all_states_years %>%
  group_by(StateName) %>%
  summarise(Average_PM2.5 = mean(Mean_PM25, na.rm = TRUE)) %>%
  arrange(desc(Average_PM2.5)) %>%
  print(n = 60)

```

## Confounder summary table

```{r}

state_confounders %>%
  select(-StateName, -Year) %>%
  select(pct_age65plus, pct_depression, pct_smoker, 
         pct_stroke, pct_lowedu, pct_lowincome, pct_nonwhite) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    label = list(
      pct_age65plus  ~ "% Age 65+",
      pct_depression ~ "% With depression",
      pct_smoker     ~ "% Current/former smoker",
      pct_stroke     ~ "% With stroke history",
      pct_lowedu     ~ "% Less than high school education",
      pct_lowincome  ~ "% Low income",
      pct_nonwhite   ~ "% Non-white"
    ),
    digits = all_continuous() ~ 1
  ) %>%
  bold_labels()

```


# Linear Mixed Models {.tabset}

Models use state as a random intercept to account for repeated measures. The outcome (mean cognitive decline) is multiplied by 100 so all beta estimates are on the percentage point scale. Each section presents unadjusted models (Year + PM2.5 lag only) alongside adjusted models (adding all confounders).

Confounders included: % age 65+, % with depression, % smokers, % with stroke history, % low education, % low income, % non-white.

## Full raw dataset {.tabset}

### Unadjusted models

```{r}

mod1_unadj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag1) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

mod2_unadj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag2) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

mod3_unadj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag3) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

tbl1 <- mod1_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- mod2_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- mod3_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Save merged table as an object
table_full_unadj <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Display it
table_full_unadj

# Export to Excel
library(openxlsx)
table_full_unadj %>%
  as_tibble() %>%
  write.xlsx("~/Desktop/Epi internship/Excel tables/table_full_unadjusted.xlsx")
```

### Adjusted models

```{r}

mod1_adj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag1) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

mod2_adj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag2) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

mod3_adj <- merged_long_data %>%
  rename(PM25_lag = PM25_lag3) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

tbl1 <- mod1_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- mod2_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- mod3_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Save merged table as an object
table_full_adj <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Display it
table_full_adj

# Export to Excel
library(openxlsx)
table_full_adj %>%
  as_tibble() %>%
  write.xlsx("~/Desktop/Epi internship/Excel tables/table_full_adjusted.xlsx")

```

### Unadjusted vs Adjusted comparison (1-year lag)

```{r}

tbl_unadj <- mod1_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy,
                 include = "PM25_lag") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_adj <- mod1_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy,
                 include = "PM25_lag") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl_unadj, tbl_adj),
  tab_spanner = c("Unadjusted", "Adjusted")
)

```


## Filtered raw dataset {.tabset}

### Unadjusted models

```{r}

mod1_unadj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag1) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

mod2_unadj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag2) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

mod3_unadj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag3) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag + (1 | StateName), data = .)

tbl1 <- mod1_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- mod2_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- mod3_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Save merged table as an object
table_filt_unadj <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Display it
table_filt_unadj

# Export to Excel
library(openxlsx)
table_filt_unadj %>%
  as_tibble() %>%
  write.xlsx("~/Desktop/Epi internship/Excel tables/table_filt_unadjusted.xlsx")

```

### Adjusted models

```{r}

mod1_adj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag1) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

mod2_adj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag2) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

mod3_adj <- filtered_merged_long_data %>%
  rename(PM25_lag = PM25_lag3) %>%
  lmer(MeanCognitiveDecline*100 ~ Year + PM25_lag
       + pct_age65plus + pct_depression + pct_smoker + pct_stroke
       + pct_lowedu + pct_lowincome + pct_nonwhite
       + (1 | StateName), data = .)

tbl1 <- mod1_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- mod2_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- mod3_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy) %>%
  add_global_p(anova_fun = function(x, ...) { anova(x, ...) |> broom::tidy() }) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Save merged table as an object
table_filt_adj <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Display it
table_filt_adj

# Export to Excel
library(openxlsx)
table_filt_adj %>%
  as_tibble() %>%
  write.xlsx("~/Desktop/Epi internship/Excel tables/table_filt_adjusted.xlsx")

```

### Unadjusted vs Adjusted comparison (1-year lag)

```{r}

tbl_unadj <- mod1_unadj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy,
                 include = "PM25_lag") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_adj <- mod1_adj %>%
  tbl_regression(pvalue_fun = ~ style_pvalue(.x, digits = 3),
                 tidy_fun = broom.mixed::tidy,
                 include = "PM25_lag") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl_unadj, tbl_adj),
  tab_spanner = c("Unadjusted", "Adjusted")
)

```


## Imputed dataset {.tabset}

Models fit on all 10 imputed datasets and pooled using Rubin's rules to account for imputation uncertainty.

### Unadjusted models

```{r}

# Rebuild mids object with original dataset included
imputed_complete_orig <- complete(imputed_data, action = "long", include = TRUE)

cog_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("MeanCognitiveDecline_"),
               names_to = "Year", names_prefix = "MeanCognitiveDecline_",
               values_to = "MeanCognitiveDecline")

pm25_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("Mean_PM25_"),
               names_to = "Year", names_prefix = "Mean_PM25_",
               values_to = "Mean_PM25")

pm25_lag1_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("PM25_lag1_"),
               names_to = "Year", names_prefix = "PM25_lag1_",
               values_to = "PM25_lag1")

pm25_lag2_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("PM25_lag2_"),
               names_to = "Year", names_prefix = "PM25_lag2_",
               values_to = "PM25_lag2")

pm25_lag3_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("PM25_lag3_"),
               names_to = "Year", names_prefix = "PM25_lag3_",
               values_to = "PM25_lag3")

conf_age_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_age65plus_"),
               names_to = "Year", names_prefix = "pct_age65plus_",
               values_to = "pct_age65plus")

conf_dep_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_depression_"),
               names_to = "Year", names_prefix = "pct_depression_",
               values_to = "pct_depression")

conf_smok_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_smoker_"),
               names_to = "Year", names_prefix = "pct_smoker_",
               values_to = "pct_smoker")

conf_stroke_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_stroke_"),
               names_to = "Year", names_prefix = "pct_stroke_",
               values_to = "pct_stroke")

conf_edu_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_lowedu_"),
               names_to = "Year", names_prefix = "pct_lowedu_",
               values_to = "pct_lowedu")

conf_inc_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_lowincome_"),
               names_to = "Year", names_prefix = "pct_lowincome_",
               values_to = "pct_lowincome")

conf_race_long <- imputed_complete_orig %>%
  pivot_longer(cols = starts_with("pct_nonwhite_"),
               names_to = "Year", names_prefix = "pct_nonwhite_",
               values_to = "pct_nonwhite")

imputed_long_orig <- cog_long %>%
  left_join(pm25_long,        by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag1_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag2_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(pm25_lag3_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_age_long,    by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_dep_long,    by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_smok_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_stroke_long, by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_edu_long,    by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_inc_long,    by = c("StateName", "Year", ".imp", ".id")) %>%
  left_join(conf_race_long,   by = c("StateName", "Year", ".imp", ".id")) %>%
  select(.imp, .id, StateName, Year, MeanCognitiveDecline, Mean_PM25,
         PM25_lag1, PM25_lag2, PM25_lag3,
         pct_age65plus, pct_depression, pct_smoker, pct_stroke,
         pct_lowedu, pct_lowincome, pct_nonwhite) %>%
  mutate(.id = row_number())

imputed_long_mids <- as.mids(imputed_long_orig)

# Fit unadjusted models
mod1_unadj <- with(imputed_long_mids,
                   lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag1 + (1 | StateName)))

mod2_unadj <- with(imputed_long_mids,
                   lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag2 + (1 | StateName)))

mod3_unadj <- with(imputed_long_mids,
                   lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag3 + (1 | StateName)))

tbl1 <- tbl_regression(mod1_unadj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- tbl_regression(mod2_unadj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- tbl_regression(mod3_unadj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

```

### Adjusted models

```{r}

mod1_adj <- with(imputed_long_mids,
                 lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag1
                      + pct_age65plus + pct_depression + pct_smoker + pct_stroke
                      + pct_lowedu + pct_lowincome + pct_nonwhite
                      + (1 | StateName)))

mod2_adj <- with(imputed_long_mids,
                 lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag2
                      + pct_age65plus + pct_depression + pct_smoker + pct_stroke
                      + pct_lowedu + pct_lowincome + pct_nonwhite
                      + (1 | StateName)))

mod3_adj <- with(imputed_long_mids,
                 lmer(MeanCognitiveDecline*100 ~ as.numeric(Year) + PM25_lag3
                      + pct_age65plus + pct_depression + pct_smoker + pct_stroke
                      + pct_lowedu + pct_lowincome + pct_nonwhite
                      + (1 | StateName)))

tbl1 <- tbl_regression(mod1_adj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl2 <- tbl_regression(mod2_adj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl3 <- tbl_regression(mod3_adj) %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)


# Save merged table as an object
table_imputed_adj <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3),
  tab_spanner = c("1-year lag", "2-year lag", "3-year lag")
)

# Display it
table_imputed_adj

# Export to Excel
library(openxlsx)
table_imputed_adj %>%
  as_tibble() %>%
  write.xlsx("~/Desktop/Epi internship/Excel tables/table_imputed_adjusted.xlsx")

```

### Unadjusted vs Adjusted comparison (1-year lag)

```{r}

tbl_unadj <- tbl_regression(mod1_unadj,
                             include = "PM25_lag1") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_adj <- tbl_regression(mod1_adj,
                           include = "PM25_lag1") %>%
  bold_p(t = 0.1) %>% bold_labels() %>% italicize_levels()

tbl_merge(
  tbls = list(tbl_unadj, tbl_adj),
  tab_spanner = c("Unadjusted", "Adjusted")
)

```
